---
title: "Project2"
author: "Amal Jain"
date: "3/9/2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
install.packages('dplyr', repos = "http://cran.us.r-project.org")
library(dplyr)
library(ggplot2)
```

1. Code for reading in the dataset and/or processing the data

```{r}
act<-read.csv("activity.csv")
act.complete <- na.omit(act)   #Creates data frame with no NA values 

```

2. Histogram of the total number of steps taken each day
```{r}
dates<-factor(act$date)
total<-tapply(act$steps,dates,sum)
hist(total, xlab = "Total daily Steps",main="Histogram of Total Steps by day",breaks=20)
```

3. Mean and median number of steps taken each day
```{r}
total<-as.data.frame.table(total)               #converts from array to data frame
colnames(total)<-c("date","total steps")

total.steps<-total[complete.cases(total),]           #Very Important
mean(total.steps$`total steps`)
median(total.steps$`total steps`)
```

4. Time series plot of the average number of steps taken
```{r}
act.interval<-group_by(act.complete,interval)  #Very useful function in dplyr library
act.interval<-summarize(act.interval,steps=mean(steps))
ggplot(act.interval, aes(interval, steps)) + geom_line()
```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
act.interval[act.interval$steps==max(act.interval$steps),]
```

6. Code to describe and show a strategy for imputing missing data
```{r}
nrow(act)-nrow(act.complete)   #no. of rows with NA vals
names(act.interval)[2] <- "mean.steps"   #Important
act.impute <- merge(act, act.interval)  #merging allows the 2 data frames to be merged together                                           on the basis of common rows and/or columns
act.impute$steps[is.na(act.impute$steps)] <- act.impute$mean.steps[is.na(act.impute$steps)]  
#upper line produces dataframe with NA values replaced with mean val of steps for that interval
```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
act.day.imp <- group_by(act.impute, date)
act.day.imp <- summarize(act.day.imp, steps=sum(steps))

qplot(steps, data=act.day.imp)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends 
```{r}
act.impute$dayofweek <- weekdays(as.Date(act.impute$date))
act.impute$weekend <-as.factor(act.impute$dayofweek=="Saturday"|act.impute$dayofweek=="Sunday")
levels(act.impute$weekend) <- c("Weekday", "Weekend")
act.weekday <- act.impute[act.impute$weekend=="Weekday",]
act.weekend <- act.impute[act.impute$weekend=="Weekend",]

act.int.weekday <- group_by(act.weekday, interval)
act.int.weekday <- summarize(act.int.weekday, steps=mean(steps))
act.int.weekday$weekend <- "Weekday"
act.int.weekend <- group_by(act.weekend, interval)
act.int.weekend <- summarize(act.int.weekend, steps=mean(steps))
act.int.weekend$weekend <- "Weekend"

act.int <- rbind(act.int.weekday, act.int.weekend)
act.int$weekend <- as.factor(act.int$weekend)
ggplot(act.int, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .)
```